Continous occlusion models and traffic scene understanding

### 2015-05-04 CVPR/Source/occlusion.tex
Latest version of presentation.

\section{Occlusion-Aware 3D Object Modeling}
\label{sec:TPmodel}

In this section, we introduce our representation of 3D traffic participants and our modeling of object-object occlusion relationships. We refer the reader to Figure [[FIGURE]] for an illustration of the proposed concepts.

\paragraph{Occupancy model for traffic participants}
Intuitively, we consider traffic participants to be regions of 3D space with a high probability of occupancy. We model the uncertainty in occupancy as a translucency function, with regions more likely to be occupied by an object considered more opaque, while regions more likely to be free space are more transparent. Based on this intuition, we model objects as translucent 3D ellipsoids whos opacity is maximum at the center and falls off towards the edges. In particular, we model the occupancy at location $\bx$ corresponding to a traffic participant $i$ centered at $\bp$ as:
\begin{align}
  \occft{\bx} = \cL(\bx; \bp, \bSigma)
\end{align}
where $\cL(\cdot)$ is the logistic function given by
\begin{align}
  L(\bx; \bp, \bSigma) = \frac{1}{1 + e^{-k(1 - d(\mathbf{x},\bp))}},
\end{align}
with $d(\bx, \bp) = (\bx-\bp)^\top\bSigma(\bx-\bp)$ being the Mahalanobis distance. We set $k = 10\ln{49}$ as the value that allows the logistic function $\cL$ to drop to $0.98$ at a distance $d = 0.9$ from the object center. The spread of the ellipsoid, determined by $\bSigma_i$, depends of the dimensions of the traffic participant. Please refer to the additional material for computation of $\bSigma$ from object dimensions.

@@

\paragraph{Reflection probability}
Consider a 3D point $\bx_j$ observed in the image at pixel $\bu_j$. Let $\ray = \displaystyle\frac{\bK^{-1}\bu_j}{\lVert \bK^{-1}\bu_j \rVert}$ be the corresponding unit vector along the backprojected ray from the camera center. Then, the probability of reflection at depth $\lambda$ along the ray $\ray$, by an object $O_i$, is determined by the object's gradient of the occupancy function $f_{occ}^i$:
\begin{align}
  P^{ij}_{\textit{reflection}} (\lambda) = (\max \{0, \nabla {f^i_{occ}}(\bx_j)^\top \ray \})^2
\end{align}
The $\max \{ \}$ ensures that the negative probability due to gradient in the direction opposite to ray is clipped off and squaring the function allows it to be smooth near zero. We note that in the extreme case of an object object, the above reverts to a (squared) Lambertian reflection.

\paragraph{Transmission probability}
A model for transmission of light through a material of thickness $\alpha$, density $\rho$ and opacity $\beta$ is given by the Beer-Lambert Law:
\begin{align}
I(x) = I_0 e^{-\beta\rho\alpha}.
\end{align}
%
In our formulation of scene occupancy, both opacity and density at a scene point $\bx_j$ are encapsulated within the total occupancy function $\occftot = \sum_i \occf$. Further, the domain of our ocupancy function $\occftot$ is $[0, 1]$ instead of $[0, \infty)$ for opacity $\beta$. Thus, we replace $e^{-\beta\rho}$ by the transparency function $1 - \occftot$ and consequently, the transmission probability over a small distance $d\lambda$ is given by
%
\begin{align}
  \!\!\!\! P_{\textit{transmission}}(\lambda + d\lambda) = P_{\textit{transmission}}(\lambda) (1-\occftot)^{d\lambda}.
\end{align}
%
For an image point $\bu_j$ to correspond to a 3D point $\bx_j$ at depth $\lambda$ along the backprojected ray $\ray$, the ray must be transmitted through space with the probability
\begin{align}
P^j_{\textit{transmission}}(\lambda) = \prod_{0}^{\lambda} (1 - \occft{\lambda \ray})^{d\lambda},
\label{eq:ptrans-integral}
\end{align}
where $\displaystyle\prod_{0}^{\lambda}$ represents the \emph{product integral} from $0$ to $\lambda$.\footnote{A product integral is a simple integral in log domain: 
\vspace{-0.2cm}
\begin{equation}
\prod_{0}^{\lambda} (1 - f_{occ}(\lambda \ray))^{d\lambda} = e^{\int_{1}^{\lambda} \ln{(1 - f_{occ}(\lambda \ray))}{d\lambda}}. \nonumber
\end{equation}
}

@@

>>>>

In this section, we introduce our representation of 3D traffic participants and our modeling of object-object occlusion relationships. We refer the reader to Figure~\ref{fig:reflectiontransimission} for an illustration of the proposed concepts.

\begin{figure}
  \usetikzlibrary{calc}
  \centering
  \begin{tikzpicture}
    \input{Source/reflectiontransmissiongraph.tex}
  \end{tikzpicture}
  \caption{We represent traffic participants as soft ellipsoids which lead to formulation of transmision and reflection probability. This figure shows the reflection probability for the first car which is highest around the camera facing face of the car. Also note that transmission probability is inversly proportional to occupancy.}
  \label{fig:reflectiontransimission}
\end{figure}

\paragraph{Occupancy model for traffic participants}
Intuitively, we consider traffic participants to be regions of 3D space with a high probability of occupancy. We model the uncertainty in occupancy as a translucency function, with regions more likely to be occupied by an object considered more opaque, while regions more likely to be free space are more transparent. Based on this intuition, we model objects as translucent 3D ellipsoids whos opacity is maximum at the center and falls off towards the edges. In particular, we model the occupancy at location $\bx$ corresponding to a traffic participant $i$ centered at $\pos{i}{t}$ as:
\begin{align}
  f_{\text{occ}}{\bx} = \cL(\bx; \pos{i}{t}, \bSigma_i)
\end{align}
where $\cL(\cdot)$ is the logistic function given by
\begin{align}
  L(\bx; \bp, \bSigma) = \frac{1}{1 + e^{-k(1 - d(\mathbf{x},\bp))}},
\end{align}
with $d(\bx, \bp) = (\bx-\bp)^\top\bSigma(\bx-\bp)$ being the Mahalanobis distance. We set $k = 10\ln{49}$ as the value that allows the logistic function $\cL$ to drop to $0.98$ at a distance $d = 0.9$ from the object center. The spread of the ellipsoid, determined by $\bSigma_i$, depends of the dimensions of the traffic participant. Please refer to the Appendix~\ref{sec:appendix} for computation of $\bSigma$ from object dimensions.

@@

\paragraph{Reflection probability}
Consider a 3D point $\bx_j$ observed in the image at pixel $\bu_j$. Let the image was formed using camera with intrinsic matrix $\bK$. Let $\ray = \frac{\bK^{-1}\bu_j}{\lVert \bK^{-1}\bu_j \rVert}$ be the corresponding unit vector along the backprojected ray from the camera center. Then, the probability of reflection at depth $\lambda$ along the ray $\ray$, by an object $O_i$, is determined by the object's gradient of the occupancy function $f_{occ}^i$:
\begin{align}
  P^{ij}_{\textit{reflection}} (\lambda) = (\max \{0, \nabla {f^i_{occ}}(\bx_j)^\top \ray \})^2
\end{align}
The $\max \{ \}$ ensures that the negative probability due to gradient in the direction opposite to ray is clipped off and squaring the function allows it to be smooth near zero. We note that in the extreme case of an object being lambertial, the above reverts to a (squared) Lambertian reflection.

\paragraph{Transmission probability}
Since we are modeling occupancy as transparency, we look towards optics for modeling of translucent objects. A model for transmission of light through a material of thickness $\alpha$, density $\rho$ and opacity $\beta$ is given by the Beer-Lambert Law:
\begin{align}
I(x) = I_0 e^{-\beta\rho\alpha}.
\end{align}
%
In our formulation of scene occupancy, both opacity and density at a scene point $\bx_j$ are encapsulated within the total occupancy function $\occftot = \sum_i \occf$. Further, the domain of our ocupancy function $\occftot$ is $[0, 1]$ instead of $[0, \infty)$ for opacity $\beta$. Thus, we replace $e^{-\beta\rho}$ by the transparency function $1 - \occftot$ and consequently, the transmission probability over a small distance $d\lambda$ is given by
%
\begin{align}
  \!\!\!\! P_{\textit{transmission}}(\lambda + d\lambda) = P_{\textit{transmission}}(\lambda) (1-\occftot)^{d\lambda}.
\end{align}
%
For an image point $\bu_j$ to correspond to a 3D point $\bx_j$ at depth $\lambda$ along the backprojected ray $\ray$, the ray must be transmitted through space with the probability
\begin{align}
P^j_{\textit{transmission}}(\lambda) = \prod_{f}^{\lambda} (1 - \occft{\lambda \ray})^{d\lambda},
\label{eq:ptrans-integral}
\end{align}
where $\displaystyle\prod_{f}^{\lambda}$ represents the \emph{product integral} from $f$ to $\lambda$ and $f$ is the position of camera screen which has been taken equivalent to focal length of the camera .\footnote{A product integral is a simple integral in log domain: 
\vspace{-0.2cm}
\begin{equation}
\prod_{0}^{\lambda} (1 - f_{occ}(\lambda \ray))^{d\lambda} = e^{\int_{1}^{\lambda} \ln{(1 - f_{occ}(\lambda \ray))}{d\lambda}}. \nonumber
\end{equation}
}

@@

% \paragraph{Notation}
% We summarize the notation used in the report for quick reference.
% \\
% 
% 
%   \begin{tabular}{|l|l|}
%     \hline
%     Symbol & Meaning \\
%     \hline
%     $\pos{i}{t}$ & Position of $i$th car at time $t$\\
%     $\ori{i}{t}$ & Orientation of $i$th car at time $t$\\
%      $\dimsn{i}$ & 3D bounding box of the car (dimensions)\\
%     $\state{i}{t}$ & State of car $=\{\pos{i}{t}, \ori{i}{t}, \dimsn{i}\}$\\
%     $\egop$ & Position of camera at time $t$\\
%     $\egoo$ & Orientation of camera at time $t$\\
%     $\relp{i}{t}$ & Relative car pose w.r.t. camera \\
%     $\tracklets$ & 3D points tracked on car $i$ in its own frame\\
%     $\trackp{t}$ & Projection of $\tracklets$ in camera\\
%     $\projectionOf{.}$ & Projection function for pose $\relp{i}{t}$\\
%     $\bb{i}$ & 2D bounding box of the car in image\\
%     \hline
%   \end{tabular}

### 2015-05-04 CVPR/Source/setup.tex
Added poster and a lots of tikz figures.
Latest version of presentation.

\section{Overview of Problem and Our Approach}
TODO:
\vspace{4cm}

\paragraph{Notation}

\begin{table}[h]
  \begin{tabular}{|l|l|}
    \hline
    Symbol & Meaning \\
    \hline
    $\pos{i}{t}$ & Position of $i$th car at time $t$\\
    $\ori{i}{t}$ & Orientation of $i$th car at time $t$\\
     $\dimsn{i}$ & 3D bounding box of the car (dimensions)\\
    $\state{i}{t}$ & State of car $=\{\pos{i}{t}, \ori{i}{t}, \dimsn{i}\}$\\
    $\egop$ & Position of camera at time $t$\\
    $\egoo$ & Orientation of camera at time $t$\\
    $\relp{i}{t}$ & Relative car pose w.r.t. camera \\
    $\tracklets$ & 3D points tracked on car $i$ in its own frame\\
    $\trackp{t}$ & Projection of $\tracklets$ in camera\\
    $\projectionOf{.}$ & Projection function for pose $\relp{i}{t}$\\
    $\bb{i}$ & 2D bounding box of the car in image\\
    \hline
  \end{tabular}
\end{table}

@@

The objective is to find the most likely traffic participant (TP) state given
various evidences $\mathbb{E} = \{\{\trackp{t}\}, \{\bb{i}\}, \text{lane det.},
\text{map}, \text{GPS}\}$.

Mathematically, find:
\begin{align}
  \{\state{i}{t}\}^* &= \arg \max P(\{\state{i}{t}\} | \mathbb{E})
\end{align}

Bayes rule
\begin{align}
  P(\{\state{i}{t}\} | \mathbb{E}) &=
  \frac{1}{Z}P( \mathbb{E} | \{\state{i}{t}\})P(\{\state{i}{t}\})
\end{align}

Assume conditional independence according to graphical model in
\ref{fig:graphmodel}.

\begin{multline}
  P(\{\state{i}{t}\} | \mathbb{E}) =
  \frac{1}{Z}
  \prod_{t=s_i}^{e_i}\prod_{i,j: i \ne j}(P(\state{i}{t}, \state{j}{t}))\\
%
  \prod_{i=1}^{N}
  \prod_{t=s_i}^{e_i}
  P(\bb{i} | \state{i}{t})
  P(\trackp{t} | \state{i}{t})\\
  P(\state{i}{t} | L_m(t))
  P(\state{i}{t} | \state{i}{t-1})
  P(\state{i}{t})
\end{multline}

We can formulate similar objective function in negative log domain:
\footnote{TODO:Resolve inconsistency in summation absorption in energy terms.}
\begin{multline}
  -\log{P(\{\state{i}{t}\} | \mathbb{E})} = 
  Z' + \sum_{i,j:i\ne j} \sum_{t=s_i}^{e_i}  \WEnergyCol \\
  + \sum_{i=1}^N \sum_{t=s_i}^{e_i}
    \WEnergy{box}
  + \WEnergy{track}\\
  + \WEnergy{lane}
  + \WEnergy{dyn}
  + \WEnergy{prior}
\end{multline}

>>>>

We are given video feed from a front facing monocular camera. We assume that features like car detections, point tracks, lane detections, ground plane and ego-motion are already available. We also use GPS and Map as a given input to the system. With this available information our goal is to find position, orientation and dimensions different participants in the scene.

We model this problem in factor graph formulation. Factor graphs are bipartite graphs containing factor node and variable nodes that describe the factorization of the joint probability over the random variables in the graph. Each factor node is connected to a variable node if and only if the corresponding function depends on the corresponding random variable node.

\begin{figure}
  \centering
  \newcommand{\imagewidth}{\textwidth}
  \input{Source/scenelayoutoverlayCity0961}
  \caption{A sample road scene with the unknowns of each car modeled as random variables. 
  The relating energies are shown in Figure~\ref{fig:graphmodel}}
\end{figure}
\begin{figure}
  %\includegraphics[width=\columnwidth]{Figures/graphicalModelFrom61ConstVars.pdf}
    %\tikzset{/tikz/x=0.8cm,/tikz/y=0.8cm}
  \begin{subfigure}[b]{0.55\textwidth}
    \input{Source/graphical_model_multiple}
    \caption{Graphical model for a single frame with state of car represented
    as single node.}
  \end{subfigure}
  \begin{subfigure}[b]{0.45\textwidth}
    \tikzset{/tikz/x=0.8cm,/tikz/y=0.8cm}
    \include{Source/graphicalmodelTP}
    \caption{Factor graph dependence split up by pose and dimension}
  \end{subfigure}
\caption{Graphical model. The six numbered nodes represent the unknown state variables of each car. The shaded nodes in the graphical model are observed variables. %(1) Bounding box energy: The bounding box energy ($\Energy{bbox}$) without occlusion modeling is a unary term, but with occlusion it becomes a higher order term that affects the state of occluder as well. The bounding box detection is represented by $\mathbf{d}_{bb}^i(t)$. We can see that the bounding box energy affects both occluder and the occluded. (2) Point tracks energy ($\Energy{track}$) also exhibits occluder-occluded dependency.  The available point tracks are modeled by $\mathbf{u}(t)$. (3) Collision energy ($\Energy{coll}$) should ideally be globally dependent on all traffic participants but here we represent collision among only those TP that are near enough to have a significant collision energy. (4) Orientation from lane (and map) information ($\Energy{lane}$) depends only on single traffic participant hence show in the right subfigure. (5) Dynamic energies ($\Energy{dyn}$) also affect only one traffic participant at a time but over two frames. (6) Size prior ($\Energy{size}$) is used to enforce reasonable size of traffic participants.
}
  \label{fig:graphmodel}
\end{figure}
\begin{figure}
  \centering
\end{figure}

@@

The objective is to find the most likely traffic participant (TP) state $\state{i}{t}$ given
various evidences $\mathbb{E} = \{\{\trackp{t}\}, \{\bb{i}\}, \{L_m\}\}$. This problem can be formulated as to find the states that maximize the conditional probability given the evidence,
%
\begin{align}
  \{\state{i}{t}\}^* &= \arg \max P(\{\state{i}{t}\} | \mathbb{E})\enspace.
\end{align}
%
By applying Bayes rule we get
\begin{align}
  P(\{\state{i}{t}\} | \mathbb{E}) &=
  \frac{1}{Z}P( \mathbb{E} | \{\state{i}{t}\})P(\{\state{i}{t}\}).
\end{align}
%
Assume conditional independence according to graphical model in Figure~\ref{fig:graphmodel}, we get an equivalant factorization of the joint probability distribution
%
\begin{multline}
  P(\{\state{i}{t}\} | \mathbb{E}) =
  \frac{1}{Z}
  \prod_{t=s_i}^{e_i}
  \left(
  \prod_{i,j: i \ne j}
  P(\state{i}{t}, \state{j}{t})
  P(\bb{i} | \state{i}{t}, \state{j}{t})
  P(\trackp{t} | \state{i}{t}, \state{j}{t})
\right)
\\
\left(
  \prod_{i=1}^{N}
  P(L_m(t) | \state{i}{t})
  P(\state{i}{t} | \state{i}{t-1})
  P(\state{i}{t})
\right)
  \enspace ,
\end{multline}
where $\bb{i}$ are the car detections, $\trackp{t}$ are the point tracks, $L_m(t)$ is the lane information from lane detection or GPS and Map information.
For computational reason it is common to work on factor graphs in log domain
and for historical reasons the negative log of probability is also called the energy of the system. In energy domain our factorization simplifies to
%
\begin{multline}
  -\log{P(\{\state{i}{t}\} | \mathbb{E})} = 
  -Z' 
  + \sum_{t=s_i}^{e_i}
  \left(
  \sum_{i,j:i\ne j}   
  \WEnergyCol 
   + \WpEnergy{bbox}
   + \WpEnergy{track}
\right)
  \\
  + \left(
  \sum_{i=1}^N 
  \WEnergy{lane}
  + \WEnergy{dyn}
  + \WEnergy{size}
\right)
  \enspace.
\end{multline}
%
Here $\WEnergyCol = -\log P(\state{i}{t}, \state{j}{t})$ is collision energy which penalizes localization of cars that are too close together, $\WpEnergy{bbox} = -\log P(\bb{i} | \state{i}{t}, \state{j}{t})$ is bounding box energy that encourages the alignment of projected bounding box with the detected bounding box, $\WpEnergy{track} = -\log P(\trackp{t} | \state{i}{t}, \state{j}{t})$ is the point tracks energy which encourages the alignment of re-projected point tracks with ones that are detected, $\WEnergy{lane} = - \log P(L_m(t) | \state{i}{t})$ is lane energy which encourages cars to be along the direction of lanes, $\WEnergy{dyn} = - \log P( \state{i}{t} | \state{i}{t-1} ) $ is dynamic energy which captures the motion model of the cars and finally $\WEnergy{size} = - \log P (\state{i}{t})$ is the size prior over the cars. $\lambda_{\text{\{col,bbox,track,lane,dyn,size\}}}$ are unknown parameters that weight different energies. For now we set them empirically, but eventually they need to be learned from the data.
